---
title: "Cleaning and Preparing Data: The Basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning and Preparing Data: The Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial we will work through the basic steps of loading and cleaning a
single data file that represents one "entity". By this we mean one table of data
that describes one thing, e.g. a household, or a study participant. If you have a
file of data that describes study participants *and* the households they belong
to (most likely redundantly with data in household columns repeated), then
this is a **multi-entity situation** that we will cover later.


```{r setup}
library(tidyverse)
library(knitr)
library(study.wrangler)
```

```{r}
cat(
  paste0(
    c(
      "here is a table of cars",
      kable(head(mtcars)),
      "\nand here is a skim of the whole table",
      capture_skim(skim(mtcars))
    ),
    collapse="\n"
  )
)
```

