---
title: "Cleaning and Preparing Data: The Basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning and Preparing Data: The Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial we will work through the basic steps of loading and cleaning a
single data file that represents one "entity". By this we mean one table of data
that describes one thing, e.g. a household, or a study participant. If you have a
file of data that describes study participants *and* the households they belong
to (most likely redundantly with data in household columns repeated), then
this is a **multi-entity situation** that we will cover later.

```{r setup}
library(tidyverse)
library(knitr)
# during development
devtools::load_all()
# or
# library(study.wrangler)
```

So let's start loading a tab-delimited file of toy data. Here we get the location
of the file from the data bundled with this package. You can simply provide
a relative or absolute path to any tab-delimited data file.

```{r}

households_file_path <- system.file("extdata", "toy_example/households.tsv", package = 'study.wrangler')
households_file_path

```
Now let's load it in with one of the wrangler's main functions:

```{r}

households <- entity_from_file(households_file_path)

```

If there are no warnings or errors, we now have an "entity object": `households`.

We can use the wrangler's `inspect()` function to see what is inside it.

```{r}

inspect(households)

```

There's a lot to unpack there, but we'll come back to that 

Now inspect an individual variable in detail

```{r}

inspect_variable(households, 'Construction.material')

```

And validate it

```{r}

validate(households)

```

Aha - something is wrong. Let's follow the advice and fix it.

```{r}

households <- households %>% set_entity_name('household')
validate(households)

```

Yay - now it validates. Let's have a closer look inside:

```{r}

inspect(households)

```

You can see that some top-level information about the entity is now filled in:

```
### Entity-level metadata ###

Field                 Value      
--------------------  -----------
name                  household  
description           NA         
display_name          household  
display_name_plural   households 
parent_name           NA         
```

And some basic row count stats:

```
### Row counts ###

Type                                 Count
----------------------------------  ------
Total                                    3
Complete data (no missing values)        3
```

And that the ID column table has assigned the 'Household.Id' column to the
correct entity (this one, level=0). Previously it was NA.

```
### ID columns ###

variable       entity_name    entity_level
-------------  ------------  -------------
Household.Id   household                 0
```

What else can we see from the `inspect(households)` report?

Well, there's a summary of what we know about the variables (non-ID-columns):

```
### Key variable metadata ###

variable                provider_label          data_type   data_shape    display_name   stable_id 
----------------------  ----------------------  ----------  ------------  -------------  ----------
Number.of.animals       Number of animals       integer     continuous    NA             NA        
Owns.property           Owns property           string      categorical   NA             NA        
Enrollment.date         Enrollment date         date        continuous    NA             NA        
Construction.material   Construction material   string      categorical   NA             NA        
```

The `provider_label` is the original column heading in the input file. From now
on we will refer to variables with their R-friendly names, in the `variable`
column above.

The wrangler has done a good job at auto-detecting the integer, date and string types.
We will explain how to correct erroneously inferred types later.

We'll come back to `data_shape` later too.

The next part of the report gently slaps our wrists for not providing display_names
and definitions (which make the variables more user-friendly in the UI)

```
### Variable annotation summary ###

Type                         Count
--------------------------  ------
Total number of variables        4
display_name provided*           0
definition provided              0
~~~~
* use `set_variable_display_names_from_provider_labels()` to use original column headings as-is.
```

Let's fix that according to the guidance, and also add a definition using the
generic `set_variable_metadata()` function. We can chain these together using the
magrittr pipe `%>%` when both functions operate on and return entity objects. 

```{r}

households <- households %>%
  set_variable_display_names_from_provider_labels() %>%
  set_variable_metadata('Owns.property', definition="The occupant owns the property either outright or via a mortgage")

inspect(households)

```

Yay, now we have `display_names` for all the variables and the second table shows one definition provided.

```
### Key variable metadata ###

variable                provider_label          data_type   data_shape    display_name            stable_id 
----------------------  ----------------------  ----------  ------------  ----------------------  ----------
Number.of.animals       Number of animals       integer     continuous    Number of animals       NA        
Owns.property           Owns property           string      categorical   Owns property           NA        
Enrollment.date         Enrollment date         date        continuous    Enrollment date         NA        
Construction.material   Construction material   string      categorical   Construction material   NA        
~~~~
Use `inspect(entity, 'variable.name')` for more detail on individual variables


### Variable annotation summary ###

Type                         Count
--------------------------  ------
Total number of variables        4
display_name provided*           4
definition provided              1
```

You can inspect annotations like `definition` with `inspect(households, 'Owns.property')` but we won't run that again.

Maybe we need a convenience function `show_variables_missing_metadata(entity, 'definition')`?

For now we can do:

```{r}

households %>% get_variable_metadata() %>% filter(is.na(definition)) %>% pull(variable) %>% cat(sep="\n")

```

